Tail Indices and Scale Parameters in Financial Time Series

### 2017-03-28 papers/FX/draft.tex
Added the result of Hoga test.

%% \newtheorem{mdef}{Definition}

\newcommand{\eqsplit}[2]{
  \begin{equation}\label{#2}
    \begin{split}
      #1
    \end{split}
  \end{equation}}
\newcommand{\eqnsplit}[1]{
  \begin{eqnarray*}
    #1
  \end{eqnarray*}}
\newcommand{\tran}[1]{
  \tilde{#1}
}
\newcommand{\td}[2]{
  \frac{d #1}{d #2}
}
\newcommand{\pd}[2]{
  \frac{\partial #1}{\partial #2}
}
\newcommand{\ppd}[2]{
  \frac{\partial^2 #1}{\partial #2^2}
}
\newcommand{\pdd}[3]{
  \frac{\partial^2 #1}{\partial #2 \partial #3}
}
\newcommand{\otd}[1]{
  \frac{d}{d #1}
}
\newcommand{\opd}[1]{
  \frac{\partial}{\partial #1}
}
\newcommand{\oppd}[1]{
  \frac{\partial^2}{\partial #1^2}
}
\newcommand{\opdd}[2]{
  \frac{\partial^2}{\partial #1 \partial #2}
}
\newcommand{\ket}[1]{
  |#1\rangle
}
\newcommand{\bra}[1]{
  \langle#1|
}
\newcommand{\inn}[1]{
  \langle#1\rangle
}
\newcommand{\mean}[1]{
  \langle#1\rangle
}
\newcommand{\tr}{
  \text{tr}\,
}
\newcommand{\re}{
  \text{Re}\,
}
\newcommand\im{
  \text{Im}\,
}
\newcommand{\arcsinh}{
  \sinh^{-1}
}
\newcommand{\arccosh}{
  \cosh^{-1}
}
\newcommand{\erfc}{
  \text{erfc}
}
\newcommand{\E}{
  \mathbb{E}
}
\renewcommand{\P}{
  \mathbb{P}
}
\newcommand{\I}{
  \mathbf{1}
}
\newcommand{\1}[1]{
  \mathbf{1}_{\{#1\}}
}
\newcommand{\diag}{
  \text{diag\,}
}
\newcommand{\pop}[1]{
\varepsilon_{#1}
}

\newcommand{\M}{
  {\text{max}}
}
\newcommand{\m}{
  {\text{min}}
}
% \newcommand{\ph}{
%   {\text{arg}\,}
% }
\newcommand\erf{
  \text{erf}
}
\renewcommand\vec[1]{
  \mathbf{#1}
}
\newcommand\cov{
  \text{cov}
}
\newcommand\var{
  \text{var}
}
\newcommand\cor{
  \text{cor}
}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}
\newcommand\mtx[1]{
  \mathbf{#1}
}
\renewcommand \L {
  \mathfrak{leb}
}
\newcommand{\hot}{
  \text{h.o.t.}
}

%% \newtheorem{proposition}{Proposition}
\newtheorem{theorem}{Theorem}
\newtheorem{acknowledgement}{Acknowledgement}
\newtheorem{algorithm}{Algorithm}
\newtheorem{axiom}{Axiom}
\newtheorem{case}{Case}
\newtheorem{claim}{Claim}
\newtheorem{conclusion}{Conclusion}
\newtheorem{condition}{Condition}
\newtheorem{conjecture}{Conjecture}
\newtheorem{corollary}{Corollary}
\newtheorem{criterion}{Criterion}
\newtheorem{definition}{Definition}
\newtheorem{example}{Example}
\newtheorem{exercise}{Exercise}
\newtheorem{lemma}{Lemma}
\newtheorem{notation}{Notation}
\newtheorem{problem}{Problem}
\newtheorem{proposition}{Proposition}
\newtheorem{remark}{Remark}
\newtheorem{solution}{Solution}
\newtheorem{summary}{Summary}

@@

Figure \ref{fig:Energy_lower},
\ref{fig:Consumer_Staples_lower},
and \ref{fig:Information_Technology_lower}
show the Hill estimates of the
lower tail indices of the ``energy'', ``consumer staples'' and ``information
technology'' sectors of the S\&P 500 index, respectively.
\begin{figure}[htb!]
  \centering
  \begin{subfigure}[b]{\linewidth}
    \includegraphics[width=\textwidth, trim={0, 0.8cm, 0, 2cm}, clip]
                    {Energy_lower.pdf}
    \caption{``Energy'' sector of S\&P 500}
    \label{fig:Energy_lower}
  \end{subfigure}
  \begin{subfigure}[b]{\linewidth}
    \includegraphics[width=\textwidth, trim={0, 0.8cm, 0, 2cm}, clip]
                    {Consumer_Staples_lower.pdf}
    \caption{``Consumer Staples'' sector of S\&P 500}
    \label{fig:Consumer_Staples_lower}
  \end{subfigure}
  \begin{subfigure}[b]{\linewidth}
    \includegraphics[width=\textwidth, trim={0, 0.8cm, 0, 2cm}, clip]
                    {Information_Technology_lower.pdf}
    \caption{``Information Technology'' sector of S\&P 500}
    \label{fig:Information_Technology_lower}
  \end{subfigure}
  \caption{\small Hill estimates of lower tail indices of
    daily return series in sectors of Standard \& Poor
    500 index. Each circle coresponds to a Hill estimate; the grey
    triangles above and below it mark the 97.5\% and 2.5\% quantiles
    of its approximate normal distribution.
    The low and high red lines mark the median of the 2.5\% quantiles
    and median of the 97.5\% quantiles, respectively.
    The data are from {\it Yahoo Finance} and the labels on
    the horizontal axis are Yahoo symbols of the stocks. The data span
    from 2010-01-01 to 2015-01-01 and comprise $n=1304$
    observations. $k=50$ lower order statistics are used
    for computing $\alpha_n$. 
  }
\end{figure}

>>>>

\input{../../physics_common.tex}

@@

Figure \ref{fig:Energy_lower},
\ref{fig:Consumer_Staples_lower},
and \ref{fig:Information_Technology_lower}
show the Hill estimates of the
lower tail indices of the ``energy'', ``consumer staples'' and ``information
technology'' sectors of the S\&P 500 index, respectively.
\begin{figure}[htb!]
  \centering
  \begin{subfigure}[b]{\linewidth}
    \includegraphics[width=\textwidth, trim={0, 0.8cm, 0, 2cm}, clip]
                    {Energy_lower.pdf}
    \caption{``Energy'' sector of S\&P 500}
    \label{fig:Energy_lower}
  \end{subfigure}
  \begin{subfigure}[b]{\linewidth}
    \includegraphics[width=\textwidth, trim={0, 0.8cm, 0, 2cm}, clip]
                    {Consumer_Staples_lower.pdf}
    \caption{``Consumer Staples'' sector of S\&P 500}
    \label{fig:Consumer_Staples_lower}
  \end{subfigure}
  \begin{subfigure}[b]{\linewidth}
    \includegraphics[width=\textwidth, trim={0, 0.8cm, 0, 2cm}, clip]
                    {Information_Technology_lower.pdf}
    \caption{``Information Technology'' sector of S\&P 500}
    \label{fig:Information_Technology_lower}
  \end{subfigure}
  \caption{\small Hill estimates of lower tail indices of
    daily return series in sectors of Standard \& Poor
    500 index. Each circle coresponds to a Hill estimate; the grey
    triangles above and below it mark the 97.5\% and 2.5\% quantiles
    of its approximate normal distribution.
    The low and high red lines mark the median of the 2.5\% quantiles
    and median of the 97.5\% quantiles, respectively.
    The data are from {\it Yahoo Finance} and the labels on
    the horizontal axis are Yahoo symbols of the stocks. The data span
    from 2010-01-01 to 2015-01-01 and comprise $n=1304$
    observations. $k=50$ lower order statistics are used
    for computing $\alpha_n$. 
  }
\end{figure}
One can see that all the confidence intervals of stocks in the
``Energy'' and ``Consumer Staples'' sectors, as well as those of a
large portion of stocks in the ``Information Technology'' sector,
overlap. This naturally prompts an investigation into the extreme
quantile estimates of the same stocks.

%% Hence it is informative to test
%% the hypothesis that the tail indices of these return series are equal.
%% Hoga \cite{Hoga2017} devised tests suitable this purpose. We adopt his
%% approach but use our own implementation. The next section has more
%% details.

%% \subsection{Hill estimates of extreme quantiles}

%% By the design of Hoga's test, the null hypothesis can be
%% rejected due to changed scale parameters instead of changed tail
%% indices. Thus it is useful to apply the Hill scale estimator to the
%% same stock return series.

\subsection{Hoga's Test}
Hoga \cite{Hoga2017} proposed his test for a time series
$\{X_i\}_{i=1}^n$ with tail functions
$\bar F_i(\cdot)$, i.e. $\P(X_i > x) = \bar F_i(x)$. The null
hypothesis is
\begin{hypothesis}
  $\mathcal H_0$: For a given $p$, $\exists t^* \in [t_0, 1 - t_0]$
  such that
  \[
  \bar F_1^{-1}(p) = \bar F_2^{-1}(p) = \cdots =
  \bar F_{\floor{t^* n}}^{-1}(p) \neq
  \bar F_{\ceil{t^* n}}^{-1}(p) = \cdots = F_n^{-1}(p)
  \]
\end{hypothesis}
where $t_0 \in (0, 1/2)$. The alternative hypothesis is that no such
$t^*$ exists. Hoga proposed the following test statistics:
\[
Q_1 = \sup_{t \in [t_0, 1 - t_0]}
\left\{
  {
    \left[
      t (1 - t) \ln\left(
      \hat x_p(0, t)
      \over
      \hat x_p(t, 1)
      \right)
    \right]^2
    \over
    \int_{t_0}^t
    \left[
      s \ln \left(
      \hat x_p(0, s)
      \over
      \hat x_p(0, t)
      \right)
    \right]^2 ds
    +
    \int_{t}^{1 - t_0}
    \left[
      (1 - s) \ln \left(
      \hat x_p(s, 1)
      \over
      \hat x_p(t, 1)
      \right)
    \right]^2 ds
  }
\right\}
\]
Under the null hypothesis, Hoga showed that as $n \to \infty$,
\begin{equation}
  \label{eq:Q1_distr}
  Q_1 \overset{d}{\to}
  \sup_{t \in [t_0, 1 - t_0]} q_t
\end{equation}
where
\[
q_t = 
  {
    [W(t) - t W(1)]^2
    \over
    \int_{t_0}^t [W(s) - {s \over t}W(t)]^2 ds
    +
    \int_{t}^{1 - t_0}
    \left[
      W(1) - W(s) - {1 - s \over 1 - t}
      (W(1) - W(t))
      \right]^2 ds
  }
\]
$W(t)$ with $t \in (0, 1)$ is a standard Brownian motion.

When applied to our problem, We would like to test, first of all,
whether there is a change of tail index in each of two 
series and, if there is no change in either series, whether the two
series have the same tail index. Obviously, for either question, if
the tail index changes, the extreme quantile will change.
For the first question, Hoga's test is directly applicable; for the
second question, we concatenate the two series before applying the
test.

Figure \ref{fig:Hoga_Energy_Single} shows the test statistics of
stocks in the ``Energy'' sector of the S\&P 500
index. The null hypothesis here is that the tail index remains the
same througout the selected period of time.
\begin{figure}[htb!]
  \begin{subfigure}[b]{0.5\linewidth}
    \includegraphics[width=\textwidth]{Hoga_Energy_Single.pdf}
    \caption{``Energy''}
    \label{fig:Hoga_Energy_Single}
  \end{subfigure}
  \begin{subfigure}[b]{0.5\linewidth}
    \includegraphics[width=\textwidth]{Hoga_Consumer_Staples_Single.pdf}
    \caption{``Consumer Staples''}
    \label{fig:Hoga_consumer_Staples_Single}
  \end{subfigure}
  \caption{Hoga Test Statistics of stocks in the ``Energy'' and
    ``Consumer Staples'' sectors of
    S\&P 500. The green, blue and red lines are, respectively the
    85\%, 90\% and 95\% quantiles of the distribution of
    \eqref{eq:Q1_distr}}.
\end{figure}
